---
title: "Biostat 212a Homework 4"
subtitle: "Due Mar. 5, 2024 @ 11:59PM"
author: "Jiyin (Jenny) Zhang, UID: 606331859"
date: today
format:
  html:
    theme: cosmo
    embed-resources: true
    number-sections: true
    toc: true
    toc-depth: 4
    toc-location: left
    code-fold: false
engine: knitr
knitr:
  opts_chunk: 
    fig.align: 'center'
    # fig.width: 6
    # fig.height: 4
    message: FALSE
    cache: false
---

## ISL Exercise 8.4.3 (10pts)

Consider the Gini index, classification error, and entropy in a simple 
classification setting with two classes. Create a single plot that displays each 
of these quantities as a function of $\hat p_{m1}$. The x-axis should display 
$\hat p_{m1}$, ranging from 0 to 1, and the $y$-axis should display the value of 
the Gini index, classification error, and entropy.

Hint: In a setting with two classes, $\hat p_{m1}$ = 1 − $\hat p_{m2}$. You 
could make this plot by hand, but it will be much easier to make in R.

**Answer:**

Create a data frame with $\hat p_{m1}$ ranging from 0 to 1, and calculate the 
Gini index, classification error, and entropy. 

```{r}
library(ggplot2)
p <- seq(0, 1, 0.01)
df <- data.frame(
  p = p,
  gini = p * (1 - p) * 2,
  entropy = -(p * log(p) + (1 - p) * log(1 - p)),
  class_err = 1 - pmax(p, 1 - p)
) 
print(df, width = 200)
```

Melt the data frame for ggplot:

```{r}
library(reshape2)
df_melted <- melt(df, id.vars = "p") |>
  print(width = 200)
```

Then use `ggplot2` to plot the three quantities as a function of $\hat p_{m1}$.

```{r}
# Plot using ggplot
ggplot(df_melted, aes(x = p, y = value, color = variable), na.rm = true) +
  geom_line() +
  labs(
    title = "Gini index, classification error, and entropy",
    subtitle = "as a function of p_hat",
    x = expression(hat(p)),
    y = "Quantities",
    color = "Quantities"
  )
```

## ISL Exercise 8.4.4 (10pts)

## ISL Exercise 8.4.5 (10pts)

## ISL Lab 8.3. `Boston` data set (30pts)

Follow the machine learning workflow to train regression tree, random forest, and boosting methods for predicting `medv`. Evaluate out-of-sample performance on a test set.

## ISL Lab 8.3 `Carseats` data set (30pts)

Follow the machine learning workflow to train classification tree, random forest, and boosting methods for classifying `Sales <= 8` versus `Sales > 8`. Evaluate out-of-sample performance on a test set.
