---
title: "Biostat 212a Homework 5"
subtitle: "Due Mar 11, 2024 @ 11:59PM"
author: "Jiyin (Jenny) Zhang, UID: 606331859"
date: today
format:
  html:
    theme: cosmo
    embed-resources: true
    number-sections: true
    toc: true
    toc-depth: 4
    toc-location: left
    code-fold: false
engine: knitr
knitr:
  opts_chunk: 
    fig.align: 'center'
    # fig.width: 6
    # fig.height: 4
    message: FALSE
    cache: false
---

## ISL Exercise 9.7.1 (10pts)

This problem involves hyperplanes in two dimensions.

__(a)__ Sketch the hyperplane 1 + 3X1 − X2 = 0. Indicate the set of points for 
which $1+3X_1 −X_2 > 0$, as well as the set of points for which 
$1 + 3X_1 − X_2 < 0$.

__(b)__ On the same plot, sketch the hyperplane $−2 + X_1 + 2X_2 = 0$. Indicate 
the set of points for which $−2 + X_1 + 2X_2 > 0$, as well as the set of points 
for which $−2+X_1+2X_2 <0$.

**Answer:**

Plotting the line:

$$
1 + 3X_1 − X_2 = 0 \Leftrightarrow X_2 = 1 + 3X_1

$$
Rearranging for $X_2$ makes it clear which side of the hyperplane is which:

$$
1 + 3X_1 − X_2 > 0 \Leftrightarrow X_2 < 1 + 3X_1 \\

1 + 3X_1 − X_2 < 0 \Leftrightarrow X_2 > 1 + 3X_1

$$

Plotting another line:

$$
−2 + X_1 + 2X_2 = 0 \Leftrightarrow X_2 = 1 - \frac{1}{2}X_1

$$
Rearranging for $X_2$ makes it clear which side of the hyperplane is which:

$$
−2 + X_1 + 2X_2 > 0 \Leftrightarrow  X_2 > 1 - \frac{1}{2}X_1 \\

−2 + X_1 + 2X_2 < 0 \Leftrightarrow X_2 < 1 - \frac{1}{2}X_1

$$

```{r}
library(ggplot2)

# line in a)
df <- data.frame(x1 = -10:10, x2 = 1 + 3 * (-10:10))
# line in b)
df2 <- data.frame(x1 = -10:10, x2 = 1 - 1/2 * (-10:10))

ggplot(df, aes(x = x1, y = x2)) +
  geom_line(color = "blue") + 
  # Add math expressions
  geom_text(aes(x = 0, y = -20), label ="paste('1 + 3', italic(X[1]), ' - ', italic(X[2]), ' > 0')", color = "blue", parse = TRUE) + 
  geom_text(aes(x = 0, y = 20), label = "paste('1 + 3', italic(X[1]), ' - ', italic(X[2]), ' < 0')", color = "blue", parse = TRUE) + 
  theme_minimal() +
  labs(x = "X1", y = "X2") +
  geom_line(data = df2, aes(x = x1, y = x2), color = "red") +
  geom_text(data = df2, aes(x = -5, y = 0), label = "paste('-2 + ', italic(X[1]), ' + 2', italic(X[2]), ' < 0')", color = "red", parse = TRUE) +
  geom_text(data = df2, aes(x = 7, y = 0), label = "paste('-2 + ', italic(X[1]), ' + 2', italic(X[2]), ' > 0')", color = "red", parse = TRUE) 
  
```




## ISL Exercise 9.7.2 (10pts)

We have seen that in p = 2 dimensions, a linear decision boundary takes the form 
$\beta_0 +\beta_1X_1 +\beta_2X_2 = 0$. We now investigate a non-linear decision 
boundary.

__(a)__ Sketch the curve

$$
(1+X_1)^2 + (2−X_2)^2 = 4
$$

**Answer:**

```{r}
library(ggplot2)
library(dplyr)
# install.packages("ggforce")
library(ggforce)
# install.packages("latex2exp")
library(latex2exp)
ggplot() + 
  geom_circle(data = data.frame(X1 = -1, X2 = 2, r = 2), 
              aes(x0 = X1, y0 = X2, r = r), col = "mediumseagreen", size = 1) + 
  scale_x_continuous(expand = c(0.01,0.01), limits = c(-3.5, 1.5)) + 
  scale_y_continuous(expand = c(0.01,0.01), limits = c(-0.5, 4.5)) + 
  labs(title = TeX(r'(Curve Plot: $(1 + X_1)^2 + (2 - X_2)^2 = 4$)'), 
       x = TeX(r'($X_1$)'), 
       y = TeX(r'($X_2$)')) + 
  coord_fixed() +
  geom_point(data = data.frame(X1 = -1, X2 = 2), aes(x = X1, y = X2), col = "red", size = 3)

```


__(b)__ On your sketch, indicate the set of points for which

$$
(1+X_1)^2 + (2−X_2)^2 > 4

$$

as well as the set of points for which

$$
(1+X_1)^2 + (2−X_2)^2 < 4

$$

**Answer:**

```{r}

grid <- expand.grid(X1 = seq(-3.5, 1.5, length.out = 200), 
                    X2 = seq(-0.5, 4.5, length.out = 200)) %>%
  mutate(region = case_when((1 + X1)^2 + (2 - X2)^2 > 4 ~ "gt 4", 
                            TRUE ~ "le 4"))

ggplot() + 
  geom_tile(data = grid, aes(x = X1, y = X2, fill = region), alpha = 0.5) + 
  geom_circle(data = data.frame(X1 = -1, X2 = 2, r = 2), 
              aes(x0 = X1, y0 = X2, r = r), col = "mediumseagreen", size = 1) + 
  annotate("text", x = -1, y = 2, label = TeX("$(1 + X_1)^2 + (2 - X_2)^2 \\leq 4$", output = "character"),
           hjust = 0.5, size = 3, parse = TRUE, col = "red") + 
  annotate("text", x = 0.5, y = 0, label = TeX("$(1 + X_1)^2 + (2 - X_2)^2 > 4$", output = "character"),
           hjust = 0.5, size = 3, parse = TRUE, col = "blue") + 
  scale_x_continuous(expand = c(0.01,0.01), limits = c(-3.5, 1.5)) + 
  scale_y_continuous(expand = c(0.01,0.01), limits = c(-0.5, 4.5)) + 
    scale_fill_manual(values = c("lightblue", "red")) +
  theme(legend.position = "none") + 
  labs(title = TeX(r'(Curve Plot: $(1 + X_1)^2 + (2 - X_2)^2 = 4$)'), 
       x = TeX(r'($X_1$)'), 
       y = TeX(r'($X_2$)')) + 
  coord_fixed()

```

__(c)__ Suppose that a classifier assigns an observation to the blue class if

$$
(1+X_1)^2 + (2−X_2)^2 > 4

$$

and to the red class otherwise. To what class is the observation $(0, 0)$ 
classified? $(-1, 1)$? $(2, 2)$? $(3, 8)$?

**Answer:**


```{r}
new_obs <- data.frame(X1 = c(0, -1, 2, 3), X2 = c(0, 1, 2, 8)) %>%
  mutate(region = case_when((1 + X1)^2 + (2 - X2)^2 > 4 ~ "gt 4", 
                            TRUE ~ "le 4"))


grid <- expand.grid(X1 = seq(-3.5, 3.5, length.out = 200), 
                    X2 = seq(-0.5, 8.5, length.out = 200)) %>%
  mutate(region = case_when((1 + X1)^2 + (2 - X2)^2 > 4 ~ "gt 4", 
                            TRUE ~ "le 4"))

ggplot() + 
  geom_tile(data = grid, aes(x = X1, y = X2, fill = region), alpha = 0.5) + 
  geom_circle(data = data.frame(X1 = -1, X2 = 2, r = 2), 
              aes(x0 = X1, y0 = X2, r = r), col = "mediumseagreen", size = 1) + 
  geom_point(data = new_obs, aes(x = X1, y = X2, col = region)) + 
  scale_x_continuous(expand = c(0.01,0.01), limits = c(-3.5, 3.5)) + 
  scale_y_continuous(expand = c(0.01,0.01), limits = c(-0.5, 8.5)) + 
  scale_fill_manual(values = c("lightblue", "red")) +
  scale_color_manual(values = c("blue", "red")) +
  theme(legend.position = "none") + 
  labs(title = TeX(r'(Classifier Plot: $f(X) = (1 + X_1)^2 + (2 - X_2)^2 - 4$)'), 
       x = TeX(r'($X_1$)'), 
       y = TeX(r'($X_2$)')) + 
    coord_fixed()

```
Base on the plot, we can see that:

- $(0, 0)$: The point is outside the circle, so it is classified as blue.
- $(-1, 1)$: The point is inside the circle, so it is classified as red.
- $(2, 2)$: The point is outside the circle, so it is classified as blue.
- $(3, 8)$: The point is outside the circle, so it is classified as blue.

__(d)__ Argue that while the decision boundary in (c) is not linear in terms of 
X1 and X2, it is linear in terms of $X_1$, $X_1^2$, $X_2$, and $X_2^2$.

## Support vector machines (SVMs) on the `Carseats` data set (30pts)

Follow the machine learning workflow to train support vector classifier (same as SVM with linear kernel), SVM with polynomial kernel (tune the degree and regularization parameter $C$), and SVM with radial kernel (tune the scale parameter $\gamma$ and regularization parameter $C$) for classifying `Sales<=8` versus `Sales>8`. Use the same seed as in your HW4 for the initial test/train split and compare the final test AUC and accuracy to those methods you tried in HW4.

## Bonus (10pts)

Let
$$
f(X) = \beta_0 + \beta_1 X_1 + \cdots + \beta_p X_p = \beta_0 + \beta^T X. 
$$
Then $f(X)=0$ defines a hyperplane in $\mathbb{R}^p$. Show that $f(x)$ is proportional to the signed distance of a point $x$ to the hyperplane $f(X) = 0$. 
